---
title: "Lab 3"
output: pdf_document
date: "2023-02-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#loading required packages and dataset
library(tidyverse)
library(haven)
library(ggthemes)
library(statar) 
library(tidyquant)
mto <- read_dta("mto.dta")
view(mto)



```

***Question 1***
The results from the Moving to Opportunity Experiment found a statistically significant positive relationship between higher earnings and having lived in a household that had an experimental voucher compared to a household with a Section 8 voucher and the control group for individuals who were young children in 94-98. The study found as similar positive, statistically significant relationship with college attendance and a negative, statistically significant relationship between living in an experimental voucher household and living in a high poverty % ZIP tract as an adult and single parenthood.However, the study found little evidence of higher earnings among adults who took part in the experiment in 94-98.  


```{r q2}

mto1 <- mto |> 
        filter(voucher == 0)

mean(mto1$moved)




```

The mean for `moved` for observations with voucher equal to 0 is, obviously, 0, because it makes sense that people who were in the control group did not use the experimental housing voucher.

```{r q3}

mto2 <- mto |> 
        filter(voucher == 1)

mean(mto2$moved)
```
The mean for `moved` for observations in the experimental group, otherwise known as the compliance rate, is 43%.

```{r q4}




```

